\documentclass{article}
\setlength{\headheight}{22.50113pt}
\addtolength{\topmargin}{-10.50113pt}

\input{preamble}
\input{letterfont}
\input{macros}

\fancyhead[L]{\tbo{Josh Park \\ Prof. Shkredov}}
\fancyhead[C]{\tbo{MA 45401-H01 -- Galois Theory Honors \\ Homework 6 (Mar 7)}}
\fancyhead[R]{\tbo{Spring 2025 \\ Page \thepage}}

\begin{document}

\setcounter{section}{6} % HW NUMBER
\begin{bexercise}
Find Galois groups for the following polynomials $ f $ over $ \Q $:
\end{bexercise}
\begin{enumerate}
\item $ (t\sq-3)(t\sq+1) $
  \begin{solution}
    We first note that $ t\sq - 3 $ is irreducible by Eisenstein's Criterion with $ p=3 $, and $ t\sq + 1 $ is irreducible since $ -1 $ is not a square in $ \Q $.
    Then, $ f $ has 4 roots: $ \alpha_{1,2} = \pm i $, $ \alpha_{3,4}=\pm\sqrt 3 $.
    Since $ \deg(\mu_i^\Q) = 2 $ and $ \deg(\mu_{\sqrt{3}}^\Q) = 2 $, we have that $ \fdeg{\Q(i)}{\Q}=2 $ and $ \fdeg{\Q(\sqrt 3)}{\Q} = 2 $ respectively by the tower law.
    Hence, for $ L=\Q(i,\sqrt 3) $, we have that $ \fdeg{L}{\Q} = \fdeg{L}{\Q(\sqrt 3)}\fdeg{\Q(\sqrt 3)}{\Q} = 4 $.
    Suppose we have some non-constant polynomial $ P\in L[t] $ such that $ P(\alpha_1,\alpha_2,\alpha_3,\alpha_4)=0 $.
    Let $ \sigma\in S_4 $ such that $(\alpha_{\sigma(1)},\alpha_{\sigma(2)},\alpha_{\sigma(3)},\alpha_{\sigma(4)})=0 $.
    We know that $ \sigma $ can only permute algebraic conjugates, so $ \pm i \mapsto \mp i $ and $ \pm \sqrt 3 \mapsto \mp \sqrt 3 $.
    Thus, the only options for $ \sigma $ are $ e,(12),(34),\tand (12)(34) $.
    Hence $ \Gal_\Q(f) = \{e,(12),(34),(12)(34)\}\iso \Ztw\edp\Ztw $
  \end{solution}

\item $ t^4-t^2+1 $
  \begin{solution}
    Note that $ t^4-t\sq+1 =\Phi_{12} $.
    From lecture, we saw that $ \Gal_\Q(\Phi_n) \iso \ZZ n^* $, the multiplicative group of units mod n.
    Noting that $ \ZZ{12}^*=\{1,5,7,11\} $, we can see that the order of each element is 2.
    Thus $ \Gal_\Q(\Phi_n) \iso \ZZ{12}^* \iso \Ztw\edp\Ztw $.
  \end{solution}

\item $ t^4-2 $
  \begin{solution}
    By Eisenstein's criterion with $ p=2 $, this polynomial is irreducible and the four roots are $ \alpha_{1,2} = \pm\sqrt[4]2, \alpha_{3,4}=\pm i\sqrt[4]2 $.
    The \sfe for this polynomial is $ \Q(\sqrt[4]2,i):\Q $, so let $ L = \Q(\sqrt[4]2, i) $.
    Since $ \fdeg{L}{\Q(i)} = 4 $ and $ \fdeg{\Q(i)}{\Q}=2 $, we have that $ \fdeg{L}{\Q} = 8 $ by the tower law.
    We know any permutation of roots can only permute algebraic conjugates of the same minimum polynomial over $ \Q $.
    Since $ \mu_{\sqrt[4]{2}}^\Q=t^4-2 $, it has conjugates $ \pm\sqrt[4]2,\pm i\sqrt[4]2 $.
    Also $ \mu_i^\Q=t^2+1 $, so it has conjugates $ i,-i $.
    Define a permutation $ \sigma $ such that $ \sigma(\sqrt[4]{2})=i\alpha $ and $ \sigma(i)=i $, and let $ \tau $ be complex conjugation.
    This gives us that $ \sigma^k(\sqrt[4]2) = i^k\sqrt[4]{2} $, $ \sigma^4=e $, and $ \tau^2 = e $.
    Next, we have \begin{align*}
      \tau\circ\sigma\circ\tau(\sqrt[4]{2})=\tau\circ\sigma(\tau(\sqrt[4]{2}))=\tau\circ\sigma(\sqrt[4]{2})=\tau(i\,\sqrt[4]{2})=\tau(i)\,\tau(\sqrt[4]{2})=(-i)\sqrt[4]{2}
    \end{align*}
    and
    \begin{align*}
      \sigma^{-1}(\sqrt[4]{2})=\sigma^3(\sqrt[4]{2})= i^3\,\sqrt[4]{2} = (-i)\sqrt[4]{2}.
    \end{align*}
    Hence, $ \tau\sigma\tau = \sigma\inv $.
    The identities $ \sigma^4 = e,\tau^2 = e $, and $ \tau\sigma\tau = \sigma\inv $ indicate that the Galois group of this polynomial is isomorphic to $ D_4 $, the dihedral group of 4 points.
    Hence $ \Gal_\Q(f) \iso D_4 $.
  \end{solution}
\end{enumerate}

\stepcounter{exercise}
\begin{subexercise}
  Find $ \Gal_{\F_3\lt(t\sq\rt)}\lt(\F_3\lt(t\rt)\rt) $.
\end{subexercise}
\begin{solution}
  Let $ K = \F_3\lt(t\sq\rt)$ and $ L = \F_3\lt(t\rt) $.
  In $ K[t] $, the element $ t $ is a root of the polynomial $ x^2-t^2 = (x+t)(x-t) $.
  Any field automorphism $ \sigma\in \Gal_K(L) $ can only permute between algebraic conjugates, so the only two automorphisms are $ \sigma(t) = t $ and $ \sigma(t) = -t \equiv 2t \pmod 3 $.
  Thus $ \Gal_{\F_3\lt(t\sq\rt)}\lt(\F_3\lt(t\rt)\rt) \iso \Ztw $.
\end{solution}

\begin{subexercise}
  Find $ \Gal_{\F_2\lt(t\sq\rt)}\lt(\F_2\lt(t\rt)\rt) $.
\end{subexercise}
\begin{solution}
  Similarly to the exercise above, the element $ t $ is a root of $ x\sq-t\sq = (x+t)(x-t) $.
  However, note that $ -t \equiv t \pmod 2 $.
  Thus the only automorphism possible is $ \sigma(t) = t $, whence $ \Gal_{\F_2\lt(t\sq\rt)}\lt(\F_2\lt(t\rt)\rt) = \tsgp $, the trivial group.
\end{solution}

\stepcounter{exercise}
\begin{subexercise}
  Let $K - M - L$ be a field extension and $ L:K $ is a normal extension.
  Prove that $L:M$ is also a normal extension.
\end{subexercise}
\begin{solution}
By theorem, $ L:K $ is normal \iff L is a \sf for some $ f\in K[t] $.
By definition of $ K[t] $, $ f(t)=\sum\limits_{i=0}^n c_it^i $ for $ c_i\in K $.
However, $ K-M\imp K\sseq M \imp k\in M \ \forall k\in K \imp f\in M[t]$.
Hence all coefficients of $ f $ are contained in $ M $ and $ L $ is a \sf for some $ f\in M[t] $ \iff $ L:M $ is a normal extension.
\end{solution}

\begin{subexercise}
  Give an example of three fields $K, M, L$ such that $\fdeg{L}{K} = 4$ and $ \fdeg{M}{K} = \fdeg{L}{M} = 2 $ (hence $ K-M $ and $ M-L $ are normal extensions) but $L : K$ is not a normal extension.
\end{subexercise}
\begin{solution}
Consider the tower of fields $ \Q(\sqrt[4]2):\Q(\sqrt 2):\Q $.
Observe that $ \mu_{\sqrt 2}^\Q(x) = x^2-2 $, whence $ \fdeg{\Q(\sqrt 2)}{\Q} = 2 $.
Thus $ \Q(\sqrt 2):\Q $ is normal.
Additionally, $ \mu_{\sqrt[4] 2}^{\Q(\sqrt 2)}(x) = x^2-\sqrt 2 $, whence $ \fdeg{\Q(\sqrt[4] 2)}{\Q(\sqrt 2)} = 2 $.
Thus $ \Q(\sqrt[4] 2):\Q(\sqrt 2) $ is normal.
By definition, $ L:K $ is normal extension if every $ f\in K[t] $ that has a root in $ L $ splits over $ L $.
However upon inspecting the extension $ \Q(\sqrt[4]2):\Q $, we notice that $ \mu_{\sqrt[4]2}^\Q = x^4-2 $ has roots $ \pm\sqrt[4]2 \tand \pm i\sqrt[4]2 $.
Since $ i\sqrt[4]2 \not\in \Q(\sqrt[4]2) $, the extension is thus not normal.
\end{solution}

\begin{bexercise}
  Let $ L:K $ be a \sfe for a non-constant polynomial $ f\in K[t] $.
  Prove that $ \order{\Gal_K(L)} $ divides $ (\deg f)! $.
\end{bexercise}
\begin{solution}
  Since $ L $ is the \sfe for $ f $ over $ K $, we know that $ \Gal_K(L) = \Gal_K(f) $.
  By lemma, we have that $ \Gal_K(f)\leq S_{\deg f} $ and by Lagrange, $ \order{\Gal_K(f)} \divs \order{S_{\deg f}}=(\deg f)! $.
\end{solution}

\stepcounter{exercise}
\begin{subexercise}
Let $ f = t\cb+t+1\in \F_2[t] $. Prove that $ \Gal_{\F_2}(f) $ is isomorphic to $ \Zth $.
\end{subexercise}
\begin{solution}
We can see that $ f(1) = 1 + 1 + 1 \equiv 1 \pmod 2 \neq 0 $ and $ f(0) = 1 \neq 0 $, whence $ f $ is irreducible over $ \F_2 $.
If $ \F_2(\alpha) $ is some extension field of $ \F_2 $ where $ \alpha $ is a root of $ f $, we know $ \fdeg{\F_2(\alpha)}{\F_2}=3 $.
Moreover, an $ \F_2 $-homomorphism $ \sigma:\F_2(\alpha)\to \F_2 $ is unique to the destination of $ \alpha $, and we know $ \alpha $ can only be sent to its own algebraic conjugates, of which there are 3.
Hence, $ \order{\Gal_{\F_2}(f)}=3 $, and the only group of order 3 is $ \Zth $.
\end{solution}

\begin{subexercise}
Let $ f = t\cb+t\sq+1\in \F_2[t] $. Prove that $ \Gal_{\F_2}(f) $ is isomorphic to $ S_3 $.
\end{subexercise}
\begin{solution}
Typo?
\end{solution}
\end{document}
