\documentclass{article}
\setlength{\headheight}{22.50113pt}
\addtolength{\topmargin}{-10.50113pt}

\usepackage{preamble}
\usepackage{letterfont}
\usepackage{macros}

\fancyhead[L]{\bd{Josh Park \\ Prof. Shkredov}}
\fancyhead[C]{\bd{MA 45401-H01 -- Galois Theory Honors \\ Homework 4}}
\fancyhead[R]{\bd{Spring 2025 \\ Page \thepage}}

\begin{document}

\setcounter{section}{4}
% \setcounter{exercise}{1}
\begin{exercise}
  For each of the following polynomials, construct a splitting field \( L \) over \( \Q \) and compute the degree \( \index{L}{\Q} \)
\end{exercise}

\begin{enumerate}
\item \( t^4+7t^2+12 \)
\begin{solution}
  We notice \( f(t) = t^4+7t\sq+12 = (t^2+3)(t^2+4) \), so let \( g(t) = t^2+3 \) and \( h(t) = t^2+4 \).
  We have that \( h=(t-2i)(t+2i) \), and by the rational root test \( h \) is irreducible over \( \Q \).
  Then \( h = \mu_{2i}^\Q \) and \( \Q(i)=M:\Q \) is the \sfe~for \( h \) with degree 2.
  Next, we have \( g = (t-i\sqrt 3)(t+i\sqrt 3) \), and by Eisenstein's criterion with \( p=3 \), \( g \) is irreducible.
  Let \( L:M \) be the \sfe~for \( g \).
  We already have that \( i\in M \), so \( L=M(\sqrt 3) \) and \( \fdeg L M = 2 \).
  Thus, \( L=\Q(i,\sqrt 3):\Q \) is the \sfe~for \( f \) and by the Tower Law, \( \fdeg L \Q = \fdeg L M \fdeg M \Q = 2\cdot 2 = 4. \)
\end{solution}

\item \( t^4+t^2+12 \)
\begin{solution}
  We notice \( f(t) = t^4+t^2+12 = (t^2-3)(t^2+4) \), so let \( g(t) = t^2-3 \) and \( h(t) = t^2+4 \).
  These are the same polynomials as in part 1, but this time the roots of \( g \) do not have an imaginary factor.
  However, we note that this did not have any impact on our argument in part 1, whence \( L = \Q(i,\sqrt 3):\Q \) is again the \sfe~for \( f \) and \( \fdeg{L}{\Q} = 4 \).
\end{solution}

\item \( t^{2n}-2^n \), where \( n=3,4 \).

\begin{proof}[Solution.] \( (n=3) \)
  We have that \( f(t) = t^6 - 2^3 \).
  Then, \( t = (2^3)^{1/6} \) and \( t = \sqrt{2}\cdot \veps_6^k, \) where \( k\in\Zsi \) and \( \veps_6=\exp(i\frac{2\pi}{6})=\exp(i\frac{\pi}{3}) \).
    We know the minimum polynomial of \( \veps_6 \) over \( \Q \) is the sixth cyclotomic polynomial \( \Phi_6 \), which has degree \( \phi(6) = \phi(2)\phi(3) = 2 \).
  Thus \( \fdeg{\Q(\veps_6)}{\Q} = 2 \).
  Now, \( \sqrt 2\not\in \Q(\veps_6) \), so \( L=\Q(\sqrt 2, \veps_6)=\Q(\veps_6)(\sqrt 2) \).
  Trivially, the minimum polynomial of \( \sqrt 2 \) has degree 2, hence \( \fdeg{L}{\Q(\veps_6)} = 2 \) and by the Tower Law, \( \fdeg{L}{\Q}=\fdeg{L}{\Q(\veps_6)}\fdeg{\Q(\veps_6)}{\Q} = 2\cdot 2 = 4. \)

  \( (n=4) \) We have that \( f(t) = t^8 - 2^4 \).
  Then, \( t = (2^4)^{1/8} \) and \( t = \sqrt{2}\cdot \veps_8^k \), where \( k\in\Zei \) and \( \veps_8=\exp(i\frac{2\pi}{8})=\exp(i\frac{\pi}{4}) \).
  We know the minimum polynomial of \( \veps_8 \) over \( \Q \) is the eighth cyclotomic polynomial \( \Phi_8 \), which has degree \( \phi(8) = 2^3-2^2 = 4 \).
  Thus \( \fdeg{\Q(\veps_8)}{\Q} = 4 \).
  Now, notice \( \veps_8 = \frac{\sqrt 2}{2}+i\frac{\sqrt 2}{2} \) and \( \veps_8\inv = \frac{\sqrt 2}{2}+i\frac{\sqrt 2}{2} \).
  So, \( \veps_8 + \veps_8\inv = \frac{\sqrt 2}{2} + \frac{\sqrt 2}{2} = \sqrt 2 \in \Q(\veps_8) \).
  Thus \( L = \Q(\veps_8) \) is the \sf for \( f \) over \( \Q \) and \( \fdeg{L}{\Q} = 4 \).
\end{proof}

\item \( t^{14}-1 \)
\begin{solution}
  Obviously, \( f(t) = t^{14} - 1 \) has root \( \veps_{14}^k \) for \( k\in \ZZ{14} \), so \( L = \Q(\veps_{14}) \)
  The minimum polynomial for \( \veps_{14} \) is \( \Phi_{14} \), whence \( \fdeg{L}{\Q} = \phi(14) = \phi(7)\phi(2)=6 \).
\end{solution}
\end{enumerate}

\begin{exercise}
  Let \( K-L-M \) be a field extension and \( K-L \), \( L-M \) are algebraic extensions.
  Prove that \( K-M \) is also an algebraic extension.
\end{exercise}

\begin{solution}
Suppose \( k\in K \) with \( \mu_k^L = x^n + \ell_{n-1}x^{n-1} + \cdots + \ell_0 \) for \( \ell_i\in L \).
Then by definition, \( k \) is algebraic over \( M(\ell_0,\ldots,\ell_{n-1}) \).
By theorem, we know that for some field extension \( F_1:F_2 \), \( \alpha\in F_1 \) is algebraic over \( F_2 \) \iff \( \fdeg{F_2(\alpha)}{F_2}<\infty \).
Thus, we have that \begin{align*}
  \fdeg{M(\ell_0,\ldots,\ell_{n-1})(k)}{M(\ell_0,\ldots,\ell_{n-1})} = \fdeg{\underbrace{M(\ell_0,\ldots,\ell_{n-1},k)}_{M''}}{\underbrace{M(\ell_0,\ldots,\ell_{n-1})}_{M'}} < \infty.
\end{align*}
Using a corollary from lecture, we also know that \( \fdeg{M'}{M} < \infty \).
Then by the Tower Law, we have \( \fdeg{M''}{M} = \fdeg{M''}{M'}\fdeg{M'}{M} < \infty \).
From a result in homework 2, any finite extension is necessarily algebraic.
Thus, \( k \) is algebraic over \( M \) for arbitrary \( k \in K \) \imp \( K-M \) is algebraic.
\end{solution}

\begin{exercise}
  Let \( \alpha \) be transcendental over a field \( K\subset \C \).
  Show that \( K(\alpha) \) is not algebraically closed (hint: consider the polynomial \( t\sq-\alpha \)).
\end{exercise}

\begin{solution}
Consider the polynomial \( t\sq-\alpha \in K(\alpha) \).
Assume ad absurdum that \( f \) is reducible over \( K(\alpha) \). Then, there is some \( \beta\in K(\alpha) \) such that \( \beta\sq = \alpha \).
By definition of \( K(\alpha) \), we have that \( \beta = \frac{g(\alpha)}{h(\alpha)} \) for some \( g(t),h(t)\in K[t] \) such that \( h(\alpha) \neq 0 \).
Thus, \( \beta\sq = \frac{g(\alpha)\sq}{h(\alpha)\sq} = \alpha \) and \( g(\alpha)\sq - \alpha h(\alpha)\sq = 0 \).
\begin{claim}
  \( g(x)\sq - x h(x)\sq \) is a nontrivial polynomial in \( K[x] \)
\end{claim}
\begin{subproof}
  Assume ad absurdum that \( g(x)\sq - xh(x) \equiv 0 \) (the trivial polynomial).
  Now, let \( m = \deg (g) \) and \( n = \deg (h) \).
  By definition, \( m \) and \( n \) must be integers.
  Obviously, \( \deg(g\sq) = 2m \) and \( \deg(h\sq) = 2n \), so \( \deg(x h\sq) = 2n+1 \).
  If we let \( g(x)\sq - x h(x)\sq = 0 \), then \( g(x)\sq = xh(x)\sq \).
  However, \( \deg(g(x)\sq) = \deg(xh(x)\sq) \iff 2m = 2n+1 \), but \( 2m \) is even and \( 2n+1 \) is odd, an obvious contradiction.
  Thus \( g(x)\sq - x h(x)\sq \in K[x] \) must be nontrivial.
\end{subproof}
Notice that the claim above contradicts that \( \alpha \) is transcendental over \( K \) by definition.
Thus \( K(\alpha) \) is not algebraically closed.
\end{solution}

\begin{exercise}
  Let \( L:K \) be a splitting field extension for a non-constant polynomial \( f\in K[t] \).
  Prove that \( \index{L}{K} \) divides \( (\deg f)! \) (hint: at the very end look at some binomial coefficients).
\end{exercise}

\begin{solution}
We prove this statement by induction on \( \deg f = n \).
The base case, \( \deg f = n = 1 \) is trivial, there is nothing to prove.
Now assume we have shown \( \fdeg{L}{K}\divs (\deg f)! \) for all \( 1\leq\deg f<n \).
We have two cases, one in which \( f \) is reducible and one in which it is not.
\begin{subproof}[Case 1 (\( f \) is irreducible).]
Let \( \alpha \) be a root of \( f \).
Then \( f \) factors as \( (t-\alpha)g \) for some polynomial \( g\in K(\alpha)[t] \) such that \( \deg g = n-1 \).
Furthermore, we can see that \( L:K(\alpha) \) is a \sfe~for \( g \).
By our inductive hypothesis, \( \fdeg{L}{K(\alpha)} \) divides \( (n-1)! \).
By lemma, the irreducibility of \( f \) implies \( f=\lambda\mu_\alpha^{K} \) for some constant \( \lambda\in K \), so \( \deg f = \deg \mu_\alpha^K = n \) and \( \fdeg{K(\alpha)}{K}=n \).
Then by using the Tower Law, we can see that \( \fdeg{L}{K(\alpha)}\fdeg{K(\alpha)}{K} = \fdeg{L}{K} \) divides \( n(n-1)! = n! \).
\end{subproof}
\begin{subproof}[Case 2 (\( f \) is reducible).]
  Let \( f = gh \) for polynomials \( g,h\in K[t] \),  \( M:K \) be a \sfe~for \( g \), and \( L:M \) be a \sfe~for \( h \).
  Then we can say \( \deg g = d \) for some \( 1\leq d < n \) by def degree, hence \( \deg h = n-d \).
  By our inductive hypothesis, \( \fdeg{M}{K} \) divides \( d! \) and \( \fdeg{L}{M} \) divides \( (n-d)! \).
  Then by use of the Tower Law again, we can easily see that \( \fdeg{L}{M}\fdeg{M}{K} = \fdeg{L}{K} \) divides \( d!(n-d)! \).
  Notice that \( \frac{n!}{d!(n-d)!} \) is exactly the binomial coefficient \( \binom{n}{d} \), which we know to be an integer for integers \( n,d \).
  Thus we can say that \( d!(n-d)! \) divides \( n! \) so finally we can conclude that, \( \fdeg{L}{M} \) divides \( n! \).
\end{subproof}
Since our inductive hypothesis holds in both cases for reducible and irreducible \( f \), we may conclude that it holds for all \( \deg f \geq 1 \) and hence any non-constant polynomial \( f\in K[t] \).
\end{solution}
\end{document}
